---
title: "Week6_Project2"
author: "Jeyaraman Ramalingam"
date: "10/8/2019"
output: html_document
---
## Week 6 Project  {.tabset}

### Overview

**Dataset References**

The following datasets were considered for this Project.

**Mecklenburg County Great Schools Report** - Untidy data with most of the values in one column.

[https://www.greatschools.org/north-carolina/charlotte/charlotte_mecklenburg-school-district/schools/?gradeLevels%5B%5D=e&view=table]

**Montgomery Count Crash Report** - Contains Report of crashes across years.

[https://data.montgomerycountymd.gov/Public-Safety/Crash-Reporting-Incidents-Data/bhju-22kf/data]

**AIRBNB NYC Data** - Neighbourhood information of different types of accomodation options.

[https://www.kaggle.com/dgomonov/new-york-city-airbnb-open-data/downloads/AB_NYC_2019.csv/3]

**Approach**

Below is the standard approach followed for all datasets used in this project.

* **Step - 1** Untidy data is analyzed at Source. 
* **Step - 2** Filter any unnecessary rows such as empty/NULL rows.
* **Step - 3** Use tidyr packages to transform the datasets to interpret meaningful information.
* **Step - 4** Use dplyr packages to process strings, transformations etc.
* **Step - 5** Use ggplot2 package to plot the inference for the Analysis performed in the dataset

### Mecklenburg Schools Report 

Below are different Analysis carried out in the county schools database.

* Analysis 1 - School with More Number of Reviews
* Analysis 2 - Ratings of Schools in zip code 28210 with number of students
* Analysis 3 - Zipcode with more number of students

```{r setup,echo=FALSE}
library(tidyverse)
library(dplyr)
library(data.table)
```
```{r Meck_Data_Prep,eval=TRUE}
raw_meck_data <- data.frame(read.csv("https://raw.githubusercontent.com/jey1987/DATA607/master/Week6_Project/Mecklenburg_Elem_Data.csv"))
raw_meck_data$School <- str_trim(str_replace_all(raw_meck_data$School,"Homes for sale",""))
raw_meck_data <- raw_meck_data[!apply(is.na(raw_meck_data) | raw_meck_data == "" | raw_meck_data == " ", 1, all),]
raw_meck_data_split1 <- data.frame(raw_meck_data$School)
raw_meck_data_split1[,2]=""
raw_meck_data_split1[,3]=""
names(raw_meck_data_split1)[1]="val"
names(raw_meck_data_split1)[2]="attribute"
names(raw_meck_data_split1)[3]="id"
i=1
j=1
raw_meck_data_split1[1,2]="Date"
raw_meck_data_split1[2,2]="Rating"
raw_meck_data_split1[3,2]="School Name"
raw_meck_data_split1[4,2]="Address"

while(i<=nrow(raw_meck_data_split1))
{
  raw_meck_data_split1[i,3]=j
  if(i%%4==0 & i!=nrow(raw_meck_data_split1))
  {
    j=j+1
    raw_meck_data_split1[i+1,2]="Date"
    raw_meck_data_split1[i+2,2]="Rating"
    raw_meck_data_split1[i+3,2]="School Name"
    raw_meck_data_split1[i+4,2]="Address"
  }
  i=i+1
}

raw_meck_data_split2 <- raw_meck_data[2:7]
raw_meck_data_split2 <- raw_meck_data_split2[!apply(is.na(raw_meck_data_split2) | raw_meck_data_split2 == "" | raw_meck_data_split2 == " ", 1, all),]
raw_meck_data_split1 <- spread(raw_meck_data_split1,attribute,val)
raw_meck_data_final <- data.frame(cbind(raw_meck_data_split1,raw_meck_data_split2[!names(raw_meck_data_split2) %in% names(raw_meck_data_split1)]))
raw_meck_data_final$id <- seq(1,nrow(raw_meck_data_final))

```

**Analysis 1 - School with More Number of Reviews**

```{r Meck_Analysis_1, eval=TRUE}
raw_meck_data_plot_reviews <- raw_meck_data_final %>%
  mutate(no_of_reviews = as.numeric(ifelse(Reviews == "No reviews yet",0,str_extract_all(Reviews,"[[:digit:]]+")))) %>%
  select(School.Name,no_of_reviews) %>%
  filter(no_of_reviews > 50) %>%
  arrange(desc(no_of_reviews)) %>%
  do(head(.,n=5))

ggplot(raw_meck_data_plot_reviews, aes(x=School.Name, y=no_of_reviews)) +
geom_bar(stat="identity",fill="#992266") + coord_flip() +  theme_minimal() + theme(axis.line = element_blank(),
axis.ticks = element_blank(),
plot.title = element_text(hjust = 0.5, color = "#666666"))
```

**Analysis 2 - Rating of Schools with Number of Students**

```{r Meck_Analysis_2, eval=TRUE}
raw_meck_data_plot_no_of_stu_by_Rating <- raw_meck_data_final %>%
  mutate(no_of_stu = as.numeric(Total.students.enrolled)) %>%
  mutate(zip_code = as.numeric(str_trim(sapply(str_extract_all(Address,"[[:digit:][:digit:][:digit:][:digit:][:digit:]]+"),tail,1)))) %>%
  filter(zip_code=="28210") %>%
  select(School.Name,no_of_stu,Rating) %>%
  group_by(School.Name,Rating) %>%
  summarise(total_by_Rating = sum(no_of_stu)) 

ggplot(raw_meck_data_plot_no_of_stu_by_Rating, aes(fill=School.Name, y=total_by_Rating, x=Rating)) + geom_bar( stat="identity", position="fill") 
```

**Analysis 3 - Number of Students by Zip Code**

```{r Meck_Analysis_3, eval=TRUE}
raw_meck_data_plot_no_of_stu <- raw_meck_data_final %>%
  mutate(zip_code = as.numeric(str_trim(sapply(str_extract_all(Address,"[[:digit:][:digit:][:digit:][:digit:][:digit:]]+"),tail,1)))) %>%
  mutate(no_of_stu = as.numeric(Total.students.enrolled)) %>%
  select(zip_code,no_of_stu) %>%
  group_by(zip_code) %>%
  summarise(total_by_zip = sum(no_of_stu))

ggplot(data=raw_meck_data_plot_no_of_stu, aes(zip_code,total_by_zip))+geom_jitter()
```

### NYC BNB Report

Below are different Analysis carried out in the NYC BNB Database.

* Analysis 1 - Cheaper and Costlier Neighbourhoods
* Analysis 2 - Top 10 Neighbourhoods who provide more accomodation 
* Analysis 3 - Top 10 Host with More reviews

```{r BNB_Prep , eval=TRUE}
raw_bnb_prep <- data.table(read.csv("https://raw.githubusercontent.com/jey1987/DATA607/master/Week6_Project/AB_NYC_2019.csv")) %>%
  select(id,host_id,host_name,neighbourhood,room_type,price,number_of_reviews) 
```

**Analysis 1 - Cheaper and Costlier Neighbourhoods**

```{r BNB_Analysis_1 , eval=TRUE}
raw_bnb_cheap <- raw_bnb_prep %>%
  select(neighbourhood,price,room_type) %>%
  group_by(neighbourhood) %>%
  summarise(avg_price = round(mean(price))) %>%
  arrange(avg_price) %>%
  do(head(.,n=5))

ggplot(raw_bnb_cheap, aes(x="", y=avg_price, fill=neighbourhood)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y", start=0) +
  geom_text(aes(label = avg_price), 
            position = position_stack(vjust = 0.5)) +
  labs(x = NULL, y = NULL, fill = NULL, 
       title = "Top 5 Cheaper Neighbourhoods") +
  guides(fill = guide_legend(reverse = TRUE)) +
  scale_colour_gradientn(colours=rainbow(5)) +
  theme_classic() +
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(hjust = 0.5, color = "#666666"))

raw_bnb_costly <- raw_bnb_prep %>%
  select(neighbourhood,price,id) %>%
  group_by(neighbourhood) %>%
  summarise(avg_price = round(mean(price))) %>%
  arrange(desc(avg_price)) %>%
  do(head(.,n=5))

ggplot(raw_bnb_costly, aes(x="", y=avg_price, fill=neighbourhood)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y", start=0) +
  geom_text(aes(label = avg_price), 
            position = position_stack(vjust = 0.5)) +
  labs(x = NULL, y = NULL, fill = NULL, 
       title = "Top 5 Costlier Neighbourhoods") +
  guides(fill = guide_legend(reverse = TRUE)) +
  scale_colour_gradientn(colours=rainbow(5)) +
  theme_classic() +
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(hjust = 0.5, color = "#666666"))

```

**Analysis 2 - Top 10 Neighbourhoods who provide more accomodation**

```{r BNB_Analysis_2 , eval=TRUE}

raw_bnb_no_of_rooms <- raw_bnb_prep %>%
  select(neighbourhood,id,room_type) %>%
  group_by(neighbourhood) %>%
  summarise(no_of_rooms = n()) %>%
  arrange(desc(no_of_rooms)) %>%
  do(head(.,n=10))


ggplot(raw_bnb_no_of_rooms, aes(x=neighbourhood, y=no_of_rooms)) +
geom_bar(stat="identity",fill="#002266") +coord_flip()  +  
geom_text(aes(label=no_of_rooms), vjust=-0.3, color="Black", size=3.5) + theme_minimal() + theme(axis.line = element_blank(),
axis.ticks = element_blank(),
plot.title = element_text(hjust = 0.5, color = "#666666"))



```

**Analysis 3 - Top 10 Host with More reviews**

```{r BNB_Analysis_3 , eval=TRUE}

raw_bnb_hosts_reviews <- raw_bnb_prep %>%
  select(host_name,number_of_reviews) %>%
  group_by(host_name) %>%
  summarise(no_of_reviews = sum(number_of_reviews)) %>%
  arrange(desc(no_of_reviews)) %>%
  do(head(.,n=10))

ggplot(raw_bnb_hosts_reviews, aes(x=host_name, y=no_of_reviews)) +
geom_bar(stat="identity",fill="#992266") + coord_flip() +  
geom_text(aes(label=no_of_reviews), vjust=-0.3, color="Black", size=3.5) + theme_minimal() + theme(axis.line = element_blank(),
axis.ticks = element_blank(),
plot.title = element_text(hjust = 0.5, color = "#666666"))

```

### Montgomery Count Crash Report

Below are different Analysis carried out in the Montgomery County Crash Database.

* Analysis 1 - Number of Crashses by Year
* Analysis 2 - Number of Crashses by Mile Point Directions
* Analysis 3 - Number of Hit and Runs over months in 2019 under different light conditions

```{r crash_prep , eval=TRUE}
raw_crash_prep <- data.table(read.csv("https://raw.githubusercontent.com/jey1987/DATA607/master/Week6_Project/Crash_Reporting_-_Incidents_Data.csv")) %>%
  select(Crash.Date.Time,Report.Number,Local.Case.Number,Hit.Run,Mile.Point.Direction,Light,Traffic.Control,Driver.Substance.Abuse) %>%
  mutate(Crash.Year = format(as.Date(Crash.Date.Time, format="%d/%m/%Y"),"%Y")) %>%
  mutate(Crash.Month = format(as.Date(Crash.Date.Time, format="%d/%m/%Y"),"%m"))

```

**Analysis 1 - Number of Crashses by Year**

```{r crash_Analysis_1 , eval=TRUE}

raw_crash_by_year <- raw_crash_prep %>%
  select(Local.Case.Number,Crash.Year) %>%
  group_by(Crash.Year) %>%
  summarise(no_by_year = n())

ggplot(raw_crash_by_year, aes(x=Crash.Year, y=no_by_year)) + geom_point() + geom_text(aes(label = no_by_year)) + theme_minimal() + ggtitle("Number of Crashes by Year") 

```

**Analysis 2 - Number of Crashses by Mile Point Directions**

```{r crash_Analysis_2 , eval=TRUE}
raw_crash_by_mile <- raw_crash_prep %>%
  select(Local.Case.Number,Mile.Point.Direction) %>%
  group_by(Mile.Point.Direction) %>%
  summarise(no_by_mile = n())


ggplot(raw_crash_by_mile, aes(x="", y=no_by_mile, fill=Mile.Point.Direction)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y", start=0) +
  geom_text(aes(label = no_by_mile), 
            position = position_stack(vjust = 0.5)) +
  labs(x = NULL, y = NULL, fill = NULL, 
       title = "Crashes by Mile Point") +
  guides(fill = guide_legend(reverse = TRUE)) +
  scale_colour_gradientn(colours=rainbow(5)) +
  theme_classic() +
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(hjust = 0.5, color = "#666666")) + ggtitle("Number of Crashes by Mile Point") 



```

**Analysis 3 - Number of Hit and Runs over months in 2019 under different light conditions**

```{r crash_Analysis_3 , eval=TRUE}

raw_crash_by_hit <- raw_crash_prep %>%
  select(Local.Case.Number,Hit.Run,Crash.Month,Crash.Year,Light) %>%
  filter(Hit.Run=="Yes") %>%
  filter(Crash.Year=="2019") %>%
  group_by(Crash.Month,Crash.Year,Light) %>%
  summarise(no_by_hit = n())

ggplot(raw_crash_by_hit, aes(fill=Light, y=no_by_hit, x=Crash.Month)) + geom_bar( stat="identity", position="fill") + theme(axis.text.x = element_text(angle=90)) + xlab("Month") + ylab("Crashes") + ggtitle("Crashes in 2019 by Month and Light Condition") 


```
